---
title: "Get started with shiny.val.tools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with shiny.val.tools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

`shiny.val.tools` takes a Shiny app, parses its R source statically, and
produces **per-feature validation artifacts** for use in pharma/GxP
validation packages. It does not validate the app — it produces evidence
that someone else (a developer, validation engineer, auditor) uses to
validate.

The unit of validation is a **feature subgraph**: the closed reactive
subgraph rooted at one or more outputs. For each feature subgraph the
package emits:

1. The subgraph as an interactive `visNetwork` HTML widget.
2. The list of R packages used inside that subgraph.
3. The list of functions/methods called inside that subgraph (with
   package origin).
4. A markdown documentation stub for the developer to fill in
   (intended use, risk classification, sign-off).

## Install

`shiny.val.tools` is not yet on CRAN.

```{r}
# From GitHub
remotes::install_github("Appsilon/shiny.val.tools")
```

## A 30-second tour

Point `svt_validate()` at any Shiny app — a directory or a `.zip`
archive — and it writes artifacts to `out_dir`:

```{r}
library(shiny.val.tools)

result <- svt_validate(
  app_path = "path/to/your/shiny/app",
  out_dir  = "validation"
)

result
#> <svt_validation>
#>   out_dir:     validation
#>   features:    7
#>   modules:     2
#>   doc stubs:   9
#>   inventories: 9
#>   widgets:     9
```

Per feature/module, three files land under `out_dir`:

```
validation/
├── km_plot.html          # interactive subgraph widget
├── km_plot.md            # doc stub (intended use, risk, sign-off)
├── km_plot/
│   └── inventory.json    # function-centric inventory
├── filter_panel.html
├── filter_panel.md
├── filter_panel/
│   └── inventory.json
└── ...
```

Open any `.html` file in a browser; it is self-contained.

## The pipeline, step by step

`svt_validate()` is a one-shot wrapper around five composable steps.
You can stop after any of them and inspect intermediate state.

```{r}
library(shiny.val.tools)

parsed    <- svt_parse(app_path = "path/to/app")
graph     <- svt_build_graph(parsed)
features  <- svt_slice(graph, manifest = "features.yml")  # NULL = default
inventory <- svt_inventory(graph, features)
artifacts <- svt_render(features, inventory, out_dir = "validation")
```

Each step returns an object with informative `print()` and `summary()`
methods:

```{r}
graph
#> <svt_graph>
#>   nodes:    42
#>   edges:    71
#>   warnings: 3

summary(graph)
#> svt_graph summary
#>   files:    8
#>   imports:  12
#>   sources:  2
#>   nodes:    42
#>   edges:    71
#>   warnings: 3
#>     - SVT-W004: 1
#>     - SVT-W005: 2
```

## The default slicing rule

Without a manifest, `svt_validate()` produces **one feature per
top-level output** and per top-level side-effecting observer. This is
the right starting point for most apps: every user-visible behavior
gets an artifact.

The default rule guarantees there are no unclaimed outputs — every
behavior is captured. The cost is that grouping is per-output, not
per-user-feature.

## The manifest: grouping outputs into features

Real features often span multiple outputs. The Kaplan–Meier plot, its
summary table, and the CSV download all belong to one *feature* —
"survival analysis for the primary endpoint cohort". The manifest
(`features.yml`) lets you say so.

Generate a starter manifest from the graph:

```{r}
graph <- svt_build_graph(svt_parse("path/to/app"))
svt_manifest_template(graph, out_path = "features.yml")
```

Then fill it in:

```yaml
features:
  - name: km_plot
    intended_use: |
      Generate Kaplan–Meier survival curves for the trial primary
      endpoint cohort, stratified by treatment arm.
    risk_classification: high
    roots:
      - output: km_plot
      - output: km_summary_table
      - observer: km_export
    package_categories:
      survival: method
      dplyr:    framework
      logger:   utility
```

When `svt_validate()` finds a `features.yml` in the app root (or when
you pass `manifest = "..."` explicitly), the manifest groups outputs
into features and decorates each artifact with the declared metadata.

The full schema is in `vignette("feature-subgraphs")`.

## Where to go next

- **Why this exists, and how it relates to `reactlog` / `riskmetric`.**
  See `vignette("validation-framing")`.

- **How the reactive graph is built from R source.** See
  `vignette("graph-model")`.

- **How the manifest works and how to declare validation-not-required.**
  See `vignette("feature-subgraphs")`.

- **The per-feature function-centric inventory and the
  `inventory.json` stability contract.** See `vignette("inventory")`.

- **The visNetwork widget — colors, shapes, hover behavior.** See
  `vignette("rendering")`.

- **The full SVT-W*** warning code reference.** See
  `vignette("warning-codes")`.
